Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach

### 2021-09-21 docs/attachment-bayes-factor.tex
publish supplmental

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{book}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach},
  pdfauthor={Claudio Zandonella Callegher, Tatiana Marci, Pietro De Carli, and Gianmarco Altoè},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{amsthm}
\makeatletter
\def\thm@space@setup{%
  \thm@preskip=8pt plus 2pt minus 4pt
  \thm@postskip=\thm@preskip
}
\makeatother

@@

\title{Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Supplemental Material}
\author{Claudio Zandonella Callegher, Tatiana Marci, Pietro De Carli, and Gianmarco Altoè}
\date{2021-09-14}

@@

\hypertarget{prerequisites}{%
\chapter*{Prerequisites}\label{prerequisites}}
\addcontentsline{toc}{chapter}{Prerequisites}

trial

\hypertarget{intro}{%
\chapter{Introduction}\label{intro}}

trial code

\begin{verbatim}
##   ID ID_class externalizing_sum internalizing_sum gender age_year Avm      Avp
## 1 25        2                16                 7      F     8.42 1.0 1.000000
## 2 27        2                 3                 3      F     9.00 2.0 2.166667
## 3 29        2                 5                 5      M     9.00 3.5 2.200000
## 4 30        2                11                11      M     8.92 1.6 2.666667
## 5 31        2                 0                 1      F     9.00 2.0 2.800000
## 6 32        2                 4                12      F     8.92 3.0 2.833333
##        av_m      av_p     Anxm     Anxp     anx_m     anx_p  mother   father
## 1 1.0000000 1.0000000 1.000000 1.000000 0.8333333 1.0000000  Secure   Secure
## 2 1.0000000 1.0000000 1.000000 1.000000 1.0000000 1.0000000 Anxious   Secure
## 3 1.0000000 0.8333333 1.500000 2.400000 1.0000000 0.8333333 Fearful  Anxious
## 4 0.8333333 1.0000000 1.600000 1.666667 0.8333333 1.0000000 Anxious Avoidant
## 5 1.0000000 0.8333333 1.833333 1.333333 1.0000000 1.0000000 Anxious Avoidant
## 6 1.0000000 1.0000000 3.200000 2.833333 0.8333333 1.0000000 Anxious  Fearful
##            interaction
## 1    M_Secure_F_Secure
## 2   M_Anxious_F_Secure
## 3  M_Fearful_F_Anxious
## 4 M_Anxious_F_Avoidant
## 5 M_Anxious_F_Avoidant
## 6  M_Anxious_F_Fearful
\end{verbatim}

plot trial

\includegraphics{attachment-bayes-factor_files/figure-latex/unnamed-chunk-3-1.pdf}

table trial

@@

\caption{\label{tab:unnamed-chunk-4}Attachment styles frequencies ($n_{subj} = 847$).}
\centering
\begin{tabular}[t]{rccccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{4}{c}{\textbf{Father Attachment}} & \multicolumn{1}{c}{\textbf{ }} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\textbf{Mother Attachemnt} & \textbf{Secure} & \textbf{Anxious} & \textbf{Avoidant} & \textbf{Fearful} & \textbf{Total}\\
\midrule
Secure & 125 & 49 & 49 & 8 & 231\\
Anxious & 51 & 100 & 98 & 37 & 286\\
Avoidant & 25 & 67 & 126 & 12 & 230\\
Fearful & 5 & 14 & 38 & 43 & 100\\
Total & 206 & 230 & 311 & 100 & 847\\
\bottomrule
\end{tabular}
\end{table}

trial citation

\citep{guApproximatedAdjustedFractional2018}

troppe cit \citep{mcelreathStatisticalRethinkingBayesian2020}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{book}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach},
  pdfauthor={Claudio Zandonella Callegher, Tatiana Marci, Pietro De Carli, and Gianmarco Altoè},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{amsthm}
\makeatletter
\def\thm@space@setup{%
  \thm@preskip=8pt plus 2pt minus 4pt
  \thm@postskip=\thm@preskip
}
\makeatother

@@

\title{Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Supplemental Material}
\author{\href{https://claudiozandonella.netlify.app/}{Claudio Zandonella Callegher}, Tatiana Marci, Pietro De Carli, and Gianmarco Altoè}
\date{2021-09-21}

@@

\hypertarget{introduction}{%
\chapter*{Introduction}\label{introduction}}
\addcontentsline{toc}{chapter}{Introduction}

This is the Supplemental Material of the article ``\emph{Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach}'' {[}TODO: add cit{]}. The aim of the paper is to provide a clear and detailed description of the Bayes Factor with the encompassing prior approach considering an applied example.

\hypertarget{study-summary}{%
\section*{Study Summary}\label{study-summary}}
\addcontentsline{toc}{section}{Study Summary}

When conducting a study, researchers usually have expectations based on hypotheses or theoretical perspectives they want to evaluate. Equality and inequality constraints on the model parameters are used to formalize researchers' expectations or theoretical perspectives into the so-called ``\emph{informative hypotheses}''. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are unsuitable for testing complex informative hypotheses. An alternative approach is to use the Bayes Factor. In particular, the Bayes Factor based on the ``\emph{encompassing prior approach}'' allows researchers to easily evaluate complex informative hypotheses. In this paper we provide a clear and detailed introduction to this approach. We describe all the steps and elements involved in the formalization of informative hypotheses and computation of the Bayes Factor following the encompassing prior approach. Subsequently, we provide an application of this method to a real case scenario comparing the different theoretical perspectives regarding the role of mother attachment and father attachment on children's socio-emotional development.

\hypertarget{supplemental-material-structure}{%
\section*{Supplemental Material Structure}\label{supplemental-material-structure}}
\addcontentsline{toc}{section}{Supplemental Material Structure}

In the supplemental material we provide the complete statistical analyses of the case study regarding attachment. For more information regarding the Bayes Factor with encompassing prior approach or the formalization of theoretical perspectives regarding the role of mother attachment and father attachment on children socio-emotional development, see the main article {[}TODO: add link article{]}. The supplemental material is structure in three parts:

\hypertarget{presentation}{%
\subsubsection*{1) Presentation}\label{presentation}}
\addcontentsline{toc}{subsubsection}{1) Presentation}

In this section, we describe the aim of the study and the sample included in the study.

\begin{itemize}
\tightlist
\item
  \textbf{Chapter \ref{case-study} - Case Study.} We describe the sample included in the study and present the results of the cluster analysis to classify the different attachment patterns.
\end{itemize}

\hypertarget{exteranlizing-problems}{%
\subsubsection*{2) Exteranlizing Problems}\label{exteranlizing-problems}}
\addcontentsline{toc}{subsubsection}{2) Exteranlizing Problems}

In this section, we present the analyses evaluating the different role of mother attachment and father attachment on children's externalizing problems. We follow three different approaches.

\begin{itemize}
\tightlist
\item
  \textbf{Chapter \ref{model-choice-ext} - Models Family Choice.} We discuss the appropriate models family to to take into account data characteristics.
\item
  \textbf{Chapter \ref{nhst-ext} - NHST.} Analysis are conducted following the traditional Null Hypothesis Significance Testing (NHST).
\item
  \textbf{Chapter \ref{model-comparison-ext} - Model Comparison.} Analysis are conducted according to the Model Comparison approach using the AIC and BIC criteria.
\item
  \textbf{Chapter \ref{BF-ext} - Bayes Factor.} Analysis are conducted following the Bayes Factor with encompassing prior approach.
\item
  \textbf{Chapter \ref{conclusion-ext} - Conclusion.} Results of the three approaches are discussed.
\end{itemize}

\hypertarget{interanlizing-problems}{%
\subsubsection*{3) Interanlizing Problems}\label{interanlizing-problems}}
\addcontentsline{toc}{subsubsection}{3) Interanlizing Problems}

In this section, we present the analyses evaluating the different role of mother attachment and father attachment on children's internalizing problems. We follow the three different approaches as for the externalizing problems analysis. This section is shorter than previous as we focus only on the results without repeating all the descriptions of the methods.

\begin{itemize}
\tightlist
\item
  \textbf{Chapter \ref{model-choice-int} - Models Family Choice.} We discuss the appropriate models family to to take into account data characteristics.
\item
  \textbf{Chapter \ref{nhst-int} - NHST.} Analysis are conducted following the traditional Null Hypothesis Significance Testing (NHST).
\item
  \textbf{Chapter \ref{model-comparison-int} - Model Comparison.} Analysis are conducted according to the Model Comparison approach using the AIC and BIC criteria.
\item
  \textbf{Chapter \ref{BF-int} - Bayes Factor.} Analysis are conducted following the Bayes Factor with encompassing prior approach.
\item
  \textbf{Chapter \ref{conclusion-int} - Conclusion.} Results of the three approaches are discussed.
\end{itemize}

\hypertarget{github}{%
\section*{GitHub}\label{github}}
\addcontentsline{toc}{section}{GitHub}

All the materials are available at the GitHub repository \url{https://github.com/ClaudioZandonella/Attachment}.

Refer to the \texttt{README} file for further information.

\hypertarget{part-presentation}{%
\part*{Presentation}\label{part-presentation}}
\addcontentsline{toc}{part}{Presentation}

\hypertarget{case-study}{%
\chapter{Case Study}\label{case-study}}

In this chapter, first, we describe the aim of the study. Subsequently, we describe the sample included in the study. We present general information about the sample and the results of the cluster analysis to classify the different attachment patterns. Finally, we consider children's social-emotional problems.

\hypertarget{aim-of-the-study}{%
\section{Aim of the Study}\label{aim-of-the-study}}

The aim of the study is to compare the different theoretical perspectives regarding the role of mother attachment and father attachment on children's social emotional development. In the literature, four different main theoretical perspectives have been identified \citep{brethertonFathersAttachmentTheory2010}:

\begin{itemize}
\tightlist
\item
  \textbf{Monotropy Theory} - only the mother has an impact on children's development.
\item
  \textbf{Hierarchy Theory} - the mother has a greater impact on children development than the the father.
\item
  \textbf{Independence Theory} - all attachment figures are equally important but they affect the children's development differently.
\item
  \textbf{Integration Theory} - to understand the impact on children's development it is necessary to consider attachment relationships taken together.
\end{itemize}

Contrasting results have been reported by studies investigating which is the ``correct'' theory. No study, however, has tried to properly evaluate the different theoretical perspectives by directly comparing the different hypotheses.

For more information regarding Attachment theory and the formalization of theoretical perspectives regarding the role of mother attachment and father attachment on children socio-emotional development, see the main article {[}TODO: add link article{]}.

\hypertarget{the-study-sample}{%
\section{The Study Sample}\label{the-study-sample}}

To evaluate the different role of father attachment and mother attachment, we consider 854 Italian children from third to sixth school grade. Only middle-childhood children are included in the analysis (5 participants are excluded as older than 12.3 years and 2 participants are excluded are excluded as the age is missing). The total sample size included in the analysis is 847 (50.65 \% Females). Participants frequencies by grade and gender are reported in Table\textasciitilde\ref{tab:table-gender-grade}.

@@

\caption{\label{tab:table-gender-grade}Participants frequancies by grade and gender ($n_{subj} = 847$).}
\centering
\begin{tabular}[t]{>{}rccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{2}{c}{\textbf{Gender}} & \multicolumn{1}{c}{\textbf{ }} \\
\cmidrule(l{3pt}r{3pt}){2-3}
\textbf{Grade} & \textbf{Females} & \textbf{Males} & \textbf{Total}\\
\midrule
\textbf{3rd} & 133 & 127 & 260\\
\textbf{4th} & 81 & 75 & 156\\
\textbf{5th} & 36 & 32 & 68\\
\textbf{6th} & 179 & 184 & 363\\
\textbf{Total} & 429 & 418 & 847\\
\bottomrule
\end{tabular}
\end{table}

Participants age mean is 10.28 (SD = 1.34). The participants age summary statistics are reported below and the distribution is presented in Figure\textasciitilde\ref{fig:plot-age}.

\begin{verbatim}
##   # Age summary statistics
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##    7.92    9.00   10.25   10.28   11.58   12.25
\end{verbatim}

\begin{figure}

{\centering \includegraphics{attachment-bayes-factor_files/figure-latex/plot-age-1}

}

\caption{Participants age distribution ($n_{subj} = 847$)}\label{fig:plot-age}
\end{figure}

\hypertarget{attachment-styles}{%
\section{Attachment Styles}\label{attachment-styles}}

Attachment towards the mother and the father was measured separately using the brief Experiences in Close Relationships Scale (ECR; {[}TODO add references{]}) completed by the children. Four main attachment styles have been recognized in the literature according to thee levels of \emph{anxiety} and \emph{avoidance}:

\begin{itemize}
\tightlist
\item
  \textbf{Secure Attachment -} children with low levels of anxiety and avoidance.
\item
  \textbf{Anxious Attachment -} children characterized by high levels of anxiety.
\item
  \textbf{Avoidant Attachment -} children characterized by high levels of avoidance.
\item
  \textbf{Fearful Attachment -} children with high levels of anxiety and avoidance.
\end{itemize}

To identify children attachment styles towards the mother and towards the father, we conduct two separate cluster analyses. Clusters are obtained using the function \texttt{hclust()} (with argument \texttt{method="ward.D"}) considering \emph{Euclidean} distances between participants responses to the ECR items.

\hypertarget{mother-cluster-analysis}{%
\subsection{Mother Cluster Analysis}\label{mother-cluster-analysis}}

Regarding mother attachment, 4 groups are selected from the cluster analysis results (see Figure\textasciitilde\ref{fig:plot-cluster-mother}).

\begin{figure}

{\centering \includegraphics{attachment-bayes-factor_files/figure-latex/plot-cluster-mother-1}

}

\caption{Mother attachment cluster dendrogram ($n_{subj} = 847$).}\label{fig:plot-cluster-mother}
\end{figure}

In Figure\textasciitilde\ref{fig:plot-scores-mother}, Anxiety (``\emph{Anx}'') and Avoidance (``\emph{Av}'') scores are presented according to mother attachment styles. The frequencies of mother attachment styles according to gender are reported in Table\textasciitilde{}\ref{tab:table-cluster-mother}.

\begin{figure}

{\centering \includegraphics{attachment-bayes-factor_files/figure-latex/plot-scores-mother-1}

}

\caption{Anxiety (“Anx”) and Avoidance (“Av”) scores according to mother attachment styles. Dashed lines represents overall average values ($n_{subj} = 847$).}\label{fig:plot-scores-mother}
\end{figure}

\caption{\label{tab:table-cluster-mother}Mother attachment styles by gender ($n_{subj} = 847$).}
\centering
\begin{tabular}[t]{>{}rcccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{4}{c}{\textbf{Attachment Style}} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\textbf{Gender} & \textbf{Secure} & \textbf{Anxious} & \textbf{Avoidant} & \textbf{Fearful}\\
\midrule
\textbf{Females} & 132 (16\%) & 131 (15\%) & 111 (13\%) & 55 ( 6\%)\\
\textbf{Males} & 99 (12\%) & 155 (18\%) & 119 (14\%) & 45 ( 5\%)\\
\textbf{Total} & 231 (27\%) & 286 (34\%) & 230 (27\%) & 100 (12\%)\\
\bottomrule
\end{tabular}
\end{table}

Compared to the overall average values of anxiety and avoidance, we can observe that:

\begin{itemize}
\tightlist
\item
  \emph{Secure} children have lower levels of anxiety and avoidance
\item
  \emph{Anxious} children have higher levels of anxiety and about the average levels of avoidance
\item
  \emph{Avoidant} children have lower levels of anxiety and higher levels of avoidance
\item
  \emph{Fearful} children have higher levels of anxiety and avoidance
\end{itemize}

These results reflect the traditional definition of the four attachment styles. Moreover, attachment styles frequencies are in lines with general results in the literature {[}TODO: add reference values{]}.

\hypertarget{mclust-check}{%
\subsubsection*{Mclust Check}\label{mclust-check}}
\addcontentsline{toc}{subsubsection}{Mclust Check}

To evaluate if 4 cluster is an appropriate choice, we compare the BIC values of different model-based clustering options. To do that we use the function \texttt{mclustBIC()} from the \texttt{mclust} R-package \citep{scruccaMclustClusteringClassification2016}. BIC values of the best three model-based clustering are reported below and overall results are presented in Figure\textasciitilde\ref{fig:plot-mclust-mother}. See \texttt{?mclustBIC()} help page for further information.

\begin{verbatim}
##   # Best model-based clustering for mother attachment
## Best BIC values:
##              EEV,6       EEV,7      EEV,8
## BIC      -23978.93 -24842.2437 -25039.952
## BIC diff      0.00   -863.3168  -1061.025
\end{verbatim}

\begin{figure}

{\centering \includegraphics{attachment-bayes-factor_files/figure-latex/plot-mclust-mother-1}

}

\caption{BIC values of different model-based clustering options. See '?mclustBIC()' help page for further information.}\label{fig:plot-mclust-mother}
\end{figure}

Results indicate the possible presence of a larger number of clusters. However, considering only 4 clusters seems to us the most reasonable choice. This is in line with attachment theory and general results in the literature. The important thing is hat no smaller number of clusters provided better results than the division in 4 clusters.

\hypertarget{father-cluster-analysis}{%
\subsection{Father Cluster Analysis}\label{father-cluster-analysis}}

Regarding father attachment, 4 groups are selected from the cluster analysis results (see Figure\textasciitilde\ref{fig:plot-cluster-father}).

\begin{figure}

{\centering \includegraphics{attachment-bayes-factor_files/figure-latex/plot-cluster-father-1}

}

\caption{Father attachment cluster dendrogram ($n_{subj} = 847$).}\label{fig:plot-cluster-father}
\end{figure}

In Figure\textasciitilde\ref{fig:plot-scores-father}, Anxiety (``\emph{Anx}'') and Avoidance (``\emph{Av}'') scores are presented according to father attachment styles. The frequencies of father attachment styles according to gender are reported in Table\textasciitilde{}\ref{tab:table-cluster-father}.

\begin{figure}

{\centering \includegraphics{attachment-bayes-factor_files/figure-latex/plot-scores-father-1}

}

\caption{Anxiety (“Anx”) and Avoidance (“Av”) scores according to father attachment styles. Dashed lines represents overall average values ($n_{subj} = 847$).}\label{fig:plot-scores-father}
\end{figure}

\caption{\label{tab:table-cluster-father}Father attachment styles by gender ($n_{subj} = 847$).}
\centering
\begin{tabular}[t]{>{}rcccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{4}{c}{\textbf{Attachment Style}} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\textbf{Gender} & \textbf{Secure} & \textbf{Anxious} & \textbf{Avoidant} & \textbf{Fearful}\\
\midrule
\textbf{Females} & 99 (12\%) & 104 (12\%) & 163 (19\%) & 63 ( 7\%)\\
\textbf{Males} & 107 (13\%) & 126 (15\%) & 148 (17\%) & 37 ( 4\%)\\
\textbf{Total} & 206 (24\%) & 230 (27\%) & 311 (37\%) & 100 (12\%)\\
\bottomrule
\end{tabular}
\end{table}

Compared to the overall average values of anxiety and avoidance, we can observe that:

\begin{itemize}
\tightlist
\item
  \emph{Secure} children have lower levels of anxiety and avoidance
\item
  \emph{Anxious} children have lower levels of anxiety and avoidance
\item
  \emph{Avoidant} children have lower levels of anxiety and higher levels of avoidance
\item
  \emph{Fearful} children have higher levels of anxiety and avoidance
\end{itemize}

These results are no as good as in the mother attachment classification, but they are still reasonable. Moreover, attachment styles frequencies are close to the general results in the literature {[}TODO: add reference values{]}.

\hypertarget{mclust-check-1}{%
\subsubsection*{Mclust Check}\label{mclust-check-1}}
\addcontentsline{toc}{subsubsection}{Mclust Check}

To evaluate if 4 cluster is an appropriate choice, we compare the BIC values of different model-based clustering options as before. BIC values of the best three model-based clustering are reported below and overall results are presented in Figure\textasciitilde\ref{fig:plot-mclust-father}.

\begin{verbatim}
##   # Best model-based clustering for father attachment
## Best BIC values:
##              EEV,9       EEV,8      EEV,5
## BIC      -25351.43 -25790.9130 -25905.197
## BIC diff      0.00   -439.4807   -553.765
\end{verbatim}

\begin{figure}

{\centering \includegraphics{attachment-bayes-factor_files/figure-latex/plot-mclust-father-1}

}

\caption{BIC values of different model-based clustering options. See '?mclustBIC()' help page for further information.}\label{fig:plot-mclust-father}
\end{figure}

As for the mother, results indicate the possible presence of a larger number of clusters. However, considering only 4 clusters seems to us the most reasonable choice for the same reasons as before.

\hypertarget{mother-and-father}{%
\subsection{Mother and Father}\label{mother-and-father}}

The frequencies of mother attachment and father attachment styles are reported in Table\textasciitilde\ref{tab:table-cluster-mother-father}.

\caption{\label{tab:table-cluster-mother-father}Mother attachment and father attachemtn styles ($n_{subj} = 847$).}
\centering
\begin{tabular}[t]{>{}rcccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{4}{c}{\textbf{Father Attachment}} \\
\cmidrule(l{3pt}r{3pt}){2-5}
\textbf{Mother Attachment} & \textbf{Secure} & \textbf{Anxious} & \textbf{Avoidant} & \textbf{Fearful}\\
\midrule
\textbf{Secure} & 125 (15\%) & 49 ( 6\%) & 49 ( 6\%) & 8 (1\%)\\
\textbf{Anxious} & 51 ( 6\%) & 100 (12\%) & 98 (12\%) & 37 (4\%)\\
\textbf{Avoidant} & 25 ( 3\%) & 67 ( 8\%) & 126 (15\%) & 12 (1\%)\\
\textbf{Fearful} & 5 ( 1\%) & 14 ( 2\%) & 38 ( 4\%) & 43 (5\%)\\
\bottomrule
\end{tabular}
\end{table}

We can observe how values on the diagonal (same attachment styles towards both parents) tends to be larger than others.

\hypertarget{children-outcomes}{%
\section{Children Outcomes}\label{children-outcomes}}

Children's social-emotional development was measured using the Strength \& Difficulties Questionnaire (SDQ; {[}TODO add references{]}) completed by the teachers. Separate scores for externalizing and internalizing problems were obtained as sum of the questionnaire items.

\hypertarget{externalizing-problems}{%
\subsection{Externalizing Problems}\label{externalizing-problems}}

Participants externalizing problems mean is 3.35 (SD = 3.91). The participants externalizing problems summary statistics are reported below and the distribution is presented in Figure\textasciitilde\ref{fig:plot-externalizing-dist}.

\begin{verbatim}
##   # Externalizing problems summary statistics
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##   0.000   0.000   2.000   3.349   6.000  18.000
\end{verbatim}

\begin{figure}

{\centering \includegraphics{attachment-bayes-factor_files/figure-latex/plot-externalizing-dist-1}

}

\caption{Participants externalizing problems distribution ($n_{subj} = 847$)}\label{fig:plot-externalizing-dist}
\end{figure}

Overall, externalizing problems are low. This is expected as the sample is not a clinical sample. Externalizing problems according to attachment styles are reported in Table\textasciitilde\ref{tab:table-cluster-ext}.
